Tissue Enrichment Analysis: TEA

### 2016-03-21 tests/hgf_benchmarking/doc/tea_main_text.tex
Added fetch function and writing
fetch_dictionary will now fetch the tissue dictionary.

Writing.

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   email={dangeles@caltech.edu}   % email address
]{\inits{DA}\fnm{David} \snm{Angeles-Albores}}
\author[
   addressref={aff2},                   % id's of addresses, e.g. {aff1,aff2}
   email={raymond@caltech.edu}   % email address
]{\inits{RYL}\fnm{Raymond Y} \snm{Lee}}
\author[
   addressref={aff2},                   % id's of addresses, e.g. {aff1,aff2}
   email={azurebrd@caltech.edu}   % email address
]{\inits{JC}\fnm{Juancarlos} \snm{Chan}}
\author[
   addressref={aff1, aff3},
   corref={aff1},
   email={pws@caltech.edu}
]{\inits{PWS}\fnm{Paul W} \snm{Sternberg}}

@@

\address[id=aff1]{%                           % unique id
  \orgname{California Institute of Technology, Division of Biology and Biological Engineering}, % university, etc
  \street{1200 E California Blvd},                     %
  \postcode{91125},                               % post or zip code
  \city{Pasadena},                              % city
  \cny{US}                                    % country
}

\address[id=aff3]{%                           % unique id
  \orgname{HHMI}, % university, etc
}

% raymond's address?
\address[id=aff2]{%
  \orgname{California Institute of Technology},
  \street{1200 E California Blvd},                     %
  \postcode{91125},                                % post or zip code
  \city{Pasadena},                              % city
  \cny{US}                                    % country
}

@@

\begin{abstract} % abstract
\parttitle{Background} %if any
Over the last ten years, there has been an explosive development in tools
capable of measuring gene expression. These tools generate a large number
of gene targets, but understanding these datasets, and forming hypotheses
based on them remains challenging. One way to analyze these datasets is to
associate ontologies, which are controlled, hierarchical, descriptive vocabularies
with genes and to look for enrichment of specific terms. Although gene ontology
(GO) is available for \emph{C. elegans}, this ontology does not include
anatomy or physiology information.
\parttitle{Results} %if any
We have developed an enrichment analysis tool for the \emph{C. elegans} tissue
ontology, is available on the web via WormBase and is available for download 
using Python's standard pip installer. In order to cut down on verbosity, we
have come up with three straightforward filtering criteria that slim the ontology
by almost tenfold. 
\parttitle{Conclusions} %if any
Our Tissue Enrichment Analysis (TEA), which can be found at www.wormbase.org/tea, 
uses a standard hypergeometric function to test a slimmed-down \emph{C. elegans} tissue 
ontology and provides users with a text and graphic representation of the results. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
	RNA-seq and other high-throughput methods in biology have the ability to identify thousands of genes that are altered between conditions. These genes are often correlated in their biological characteristics or functions, but identifying these functions remains challenging. To interpret these long lists of genes, biologists need to abstract genes into fewer terms that are biologically relevant to form hypotheses about what is happening in the data. One such abstraction method relies on Gene Ontology (GO). GO provides a controlled set of hierarchically ordered terms in the form of an directed acyclic graph  \cite{TheGeneOntologyConsortium2000, Ontology2009, TheGeneOntologyConsortium2015} that provide detailed information about the molecular, cellular or biochemical functions of the gene among others. For a given gene list, certain software programs can query whether a particular gene is enriched  \cite{Mi2009, McLean2010, Huang2009}. One area of biological significance that GO does not include is physiology and anatomy. One way to address this shortcoming is to generate a `tissue ontology' that provides a complete anatomical description for an organism or sets of organisms, such as `tissue', `organ' or `neuronal cell', for example. Such a tissue ontology has been developed previously \cite{Lee2003}.  The C. elegans  database, Wormbase \cite{Harris2014}, maintains a carefully curated list of gene expression data from GFP-reporters. 
	Here we provide a new framework that analyses user-input list for enrichment of specific tissues. We believe that tissues are physiologically relevant units with broad, relatively well-understood functionalities amenable to hypothesis formation. As such, we believe that identification of tissues is likely to provide researchers with enough information to be able to form hypotheses about the physiological responses of an organism to a specified condition.

	Another problem frequently associated with GO analaysis is that it is often difficult to interpret due to the large number of terms associated with a given gene. There exist a number of GO analytic tools for use by the community but a shared complaint for many programs is the very large number of GO terms that are significantly associated with any given gene list.  A common tool for GO analysis, DAVID, clusters terms into broad categories that are amenable to exploration by researchers  \cite{Huang2007}, whereas PANTHER, a different software package  \cite{Mi2009, Mi2013}, attempts to solve this issue by employing a manually reduced ontology, GOslim (pers. comm. H. Yu and P. Thomas).

@@

\section*{Results}
\subsection*{Generating a Useful Dictionary}
\subsubsection*{Reducing term redundancy through a similarity metric}
As a first step to generate our tissue enrichment software, we wished to select tissue terms that were reasonably well-annotated, yet specific enough to provide insight and not redundant with other terms. We also wanted to avoid testing tissues at levels where redundancy becomes problematic. For example, several left and right neurons have at least 25 annotating genes and we may want to include them for enrichment testing. However, many left/right neuronal pairs (which are sisters in the ontology) have almost identical annotations, with at most one or two gene differences between them. We reasoned that when two tissues have almost identical annotations, we cannot have statistical confidence in differentiating between them. As a result, testing these sister tissues provides no additional information compared with testing only the parent node to these sisters. We refer to such sisters as `redundant'. To identify redundancy, we defined a similarity metric (see \emph{Methods} section and Figure~\ref{fig:simdiagram}). Our similarity metric can be used to identify sisters that have very high similarity between them; alternatively, redundant sisters could be identified if a single sister had a very high similarity score. We referred to these two scoring criteria as `avg' and `any' respectively.

\subsubsection*{Terminal branch terms and parent terms can be safely removed in an algorithmic fashion }
Another problem arises from the fact that the tissue ontology is scarcely populated at this point in time. Many nodes have 0-10 annotations, which we consider too few to accurately test. To solve this issue, we implemented a straightforward trimming algorithm. For a given terminal node, we test whether the node has more than a threshold number of annotations. If it does not, the node is removed. The next node in the branch is tested and removed recursively until a node which satisfies the condition is found. At that point, no more nodes can be removed from that branch. This is guaranteed by the structure of the ontology: Parent nodes inherit all of the annotations of all of their descendants, so the number of annotating terms monotonically increases with increasing term hierarchy (see Figure~\ref{fig:trim_ends}). In this way, we ensure that our term dictionary includes only those tissues that are considered sufficiently well annotated for statistical purposes.

@@

\subsection*{Validation of the algorithm and parameter selection}
	We wanted to select a dictionary which included enough terms to be specific beyond the largest C. elegans tissues, yet would minimize the number of spurious results and which had a good dynamic range in terms of enrichment fold-change. Selection of a dictionary based only on minimization of spurious results would result in a dictionary with a large number of annotations per tissue, and would therefore include only the major tissues. On the other hand, selecting a dictionary that can detect smaller tissues will bias us towards tissues with lesser annotations. To our knowledge there is no good method for assessing false-positive or false-negative results for annotations.To help us select an appropriate dictionary and validate our tool, we found a set of 30 gold standards based on microarray and RNA-seq literature which are believed to be enriched in specific tissues \cite{Gaudet2004a, Spencer2011, Cinar2005, Watson2008a, Pauli2006, Portman2004, Fox2007, Smith2010}. Some of these studies have been used to annotate gene expression, and so they did not constitute an independent testing set. To correct this flaw, we built a clean dictionary that specifically excluded all annotation evidence that came from these studies.

	As a first attempt to select a good dictionary, we generated all the possible combinations of dictionaries with minimal annotations of 10, 25, 50 and 100 genes and similarity cutoffs of 0.9, 0.95 and 1, using `avg' or `any' thresholding criteria for the latter (see Table ~\ref{tab:DictionarySpecs}). For these dictionaries, the number of tissues tested ranged from 21 to 460. The number of tissues was inversely correlated to the minimum annotation, as expected, and was largely insensitive to the redundancy threshold, at least in the range we explored (0.9-1). Next, we analyzed all 30 datasets using each dictionary. Because of the large number of results, instead of analyzing each set of terms individually, we pooled all results for a given dictionary into histograms. When we analyzed the distribution of significant q-values for the dictionaries, we found that the similarity threshold mattered relatively little for any dictionary. We also noticed that the `any' thresholding method resulted in tighter histograms with a mode closer to 0. For this reason, we chose the `any' method for dictionary generation. The average q-value increased with decreasing annotation cut-off (see Figure~\ref{fig:qvals}), which reflects the decreasing statistical power associated with fewer annotations per term, but we remained agnostic as to how significant the trade-off between power and term specificity is. Based on these observations, we ruled out the dictionary with the 100 gene annotation cut-off: it had the fewest terms and its q-values were not low enough to compensate the trade-off in specificity.

@@

\subsection*{A brief example}
We applied our tool to the RNA-seq datasets developed by Engelmann et al.  \cite{Engelmann2011} to gain further understanding of the biology underlying these datasets. Engelmann et al.  exposed young adult worms to 5 different pathogenic bacteria or fungi for 24 hours, after which mRNA was extracted from the worms for sequencing. We ran TEA on the genes Engelmann \emph{et al} identified as up- or down-regulated. Initially we noticed that genes that are down-regulated tend to be twice better annotated on average than genes that were up-regulated, suggesting that our understanding of the worm immune system is scarce, in spite of important advances made over the last decade. Strikingly, three out of the five samples showed enrichment of neuronal tissues or neuronal precursor tissues amongst the down-regulated genes (see \ref{fig:Dcon, fig:Hsp, fig:Smar}). A possible explanation for this might be that the infected worms are sick and the neurons are beginning to shut down; an alternative hypothesis would be that the worm is down-regulating specific neuronal pathways as a behavioural response against the pathogen. Indeed, several studies \cite{Meisel2014, Zhang2005} have provided evidence that \emph{C. elegans} uses chemosensory neurons to identify pathogens. Interestingly, one bacterium did not exhibit the same pattern of down-regulation of neuronal-associated genes. \emph{E. faecalis} showed increased expression of genes associated with neuronal tissues, hinting that \emph{E. faecalis} may have a different pathogenic profile. Up-regulated tissues, when detected, almost always included the hypodermis and excretory duct.
	Our results highlight the involvement of various \emph{C. elegans} neuronal tissues in pathogen defense and/or illness.

\section*{Discussion}
We have presented a tissue enrichment analysis tool that employs a standard hypergeometric model to test the \emph{C. elegans} tissue ontology. We have also presented the first, to our knowledge, ontology trimming algorithm. This algorithm, which is very easy to execute, places strong limits on the number of terms selected for testing. Due to the nature of all ontologies as hierarchical, acyclical graphs with term inheritance, term annotations are correlated along any given branch. This correlation reduces the benefits of including all terms for statistical analysis: for any given term along a branch, if that term passes significance, there is a high probability that many other terms along that branch will also pass significant. If the branch is enriched by random chance, error propagation along a branch means that many more false positives will follow. Thus, a researcher might be misled by the number of terms of correlated function and assign importance to this finding; the fact that the branching structure of GO amplifies false positive signals is a powerful argument for either reducing branch length or branch intracorrelation, or both. On the other hand, if a term is actually enriched, we argue that there is little benefit to presenting the user with additional terms along that branch. Instead, a user will benefit most from testing sparsely along the tree at a suitable specificity for hypothesis formation. Related terms of the same level should only be tested when there is sufficient annotation to differentiate, with statistical confidence, whether one term is enriched above the other. Our algorithm reduces branch length by identifying and removing nodes that are insufficiently annotated and parents that are likely to include sparse information.

Chikina \emph{et al}  \cite{Chikina2009} report a tissue enrichment model based on an SVM classifier that has been trained on microarray studies. SVM classifiers are powerful tools capable of great sensitivity, but they require continuous retraining as tissue expression data becomes more available. Moreover, classifiers require that data be rank-ordered by some metric, something which is not possible for some genome-wide studies. Our tool relies on an annotation dictionary that is continuously updated, does not require retraining and does not require ranked genes. To our knowledge, there are no other tissue ontology enrichment tools in \emph{C. elegans}, but similar projects exist for humans and zebrafish ~\cite{Lee2013, Prykhozhij2013}, highlighting the relevance of our tool for high-dimensionality biology.

@@

\section*{Availability of data and materials}
Our web implementation is available at https://www.wormbase.org/tea. Our software can also be downloaded using Python's pip installer via the command

@@

\section*{Competing interests}
 	The authors declare that they have no competing interests.
\section*{Author's contributions}
    DA and PWS conceived of the project; DA developed algorithm;
	RYL made intellectual contributions to the project; RYL and JC
	developed the web GUI.
\section*{Acknowledgements}
	We would like to acknowledge all members of the Sternberg
	lab for helpful discussion.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   % email={dangeles@caltech.edu}   % email address
]{\inits{DA}\fnm{David} \snm{Angeles-Albores}}
\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   % email={raymond@caltech.edu}   % email address
]{\inits{RYL}\fnm{Raymond Y} \snm{Lee}}
\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   % email={azurebrd@caltech.edu}   % email address
]{\inits{JC}\fnm{Juancarlos} \snm{Chan}}
\author[
   addressref={aff1},
   corref={aff1},
   email={pws@caltech.edu}
]{\inits{PWS}\fnm{Paul W} \snm{Sternberg}}

@@

\address[id=aff1]{%                           % unique id
  \orgname{HHMI and California Institute of Technology, Division of Biology and Biological Engineering}, % university, etc
  \street{1200 E California Blvd},                     %
  \postcode{91125},                               % post or zip code
  \city{Pasadena},                              % city
  \cny{US}                                    % country
}

@@

\begin{abstract} % abstract
\parttitle{Background} %if any
Over the last ten years, there has been an explosive development in tools
capable of measuring gene expression. These tools generate a large number
of gene targets, but understanding these datasets, and forming hypotheses
based on them remains challenging. One way to analyze these datasets is to
associate ontologies, which are controlled, hierarchical, descriptive vocabularies
with genes and to look for enrichment of specific terms. Although gene ontology
(GO) is available for   \emph{C. elegans}, this ontology does not include
anatomy or physiology information.
\parttitle{Results} %if any
We have developed an enrichment analysis tool for the   \emph{C. elegans} tissue
ontology, is available on the web via WormBase and is available for download 
using Python's standard pip installer. In order to cut down on verbosity, we
have come up with three straightforward filtering criteria that slim the ontology
by almost tenfold. 
\parttitle{Conclusions} %if any
Our Tissue Enrichment Analysis (TEA), which can be found at www.WormBase.org/tea, 
uses a standard hypergeometric function to test a slimmed-down   \emph{C. elegans} tissue 
ontology and provides users with a text and graphic representation of the results. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
	RNA-seq and other high-throughput methods in biology have the ability to identify thousands of genes that are altered between conditions. These genes are often correlated in their biological characteristics or functions, but identifying these functions remains challenging. To interpret these long lists of genes, biologists need to abstract genes into fewer terms that are biologically relevant to form hypotheses about what is happening in the data. One such abstraction method relies on Gene Ontology (GO). GO provides a controlled set of hierarchically ordered terms in the form of an directed acyclic graph  \cite{TheGeneOntologyConsortium2000, Ontology2009, TheGeneOntologyConsortium2015} that provide detailed information about the molecular, cellular or biochemical functions of the gene among others. For a given gene list, certain software programs can query whether a particular gene is enriched  \cite{Mi2009, McLean2010, Huang2009}. One area of biological significance that GO does not include is anatomy. One way to address this shortcoming is to generate a `tissue ontology' that provides a complete anatomical description for an organism or sets of organisms, such as `tissue', `organ' or `neuronal cell', for example. Such a tissue ontology has been developed previously \cite{Lee2003}.  The  \emph{C. elegans} database, WormBase \cite{Harris2014}, maintains a carefully curated list of gene expression data from GFP-reporters. 
	Here we provide a new framework that analyses user-input list for enrichment of specific tissues. We believe that tissues are physiologically relevant units with broad, relatively well-understood functionalities amenable to hypothesis formation. As such, we believe that identification of tissues is likely to provide researchers with enough information to be able to form hypotheses about the physiological responses of an organism to a specified condition.

	Another problem frequently associated with GO enrichment analysis is that it is often difficult to interpret due to the large number of terms associated with a given gene. There exist a number of GO analytic tools for use by the community but a shared complaint for many programs is the very large number of GO terms that are significantly associated with any given gene list.  A common tool for GO analysis, DAVID, clusters terms into broad categories that are amenable to exploration by researchers  \cite{Huang2007}, whereas PANTHER, a different software package  \cite{Mi2009, Mi2013}, attempts to solve this issue by employing a manually reduced ontology, GOslim (H. Yu and P. Thomas, pers. comm. ).

@@

\section*{Results}
\subsection*{Generating a Useful Dictionary}
\subsubsection*{Reducing term redundancy through a similarity metric}
As a first step to generate our tissue enrichment software, we wished to select tissue terms that were reasonably well-annotated, yet specific enough to provide insight and not redundant with other terms. We also wanted to avoid testing tissues at levels where redundancy becomes problematic. For example, several left and right neurons have at least 25 annotated genes and we may want to include them for enrichment testing. However, many left/right neuronal pairs (which are sisters in the ontology) have almost identical annotations, with at most one or two gene differences between them. We reasoned that when two tissues have almost identical annotations, we cannot have statistical confidence in differentiating between them. As a result, testing these sister tissues provides no additional information compared with testing only the parent node to these sisters. We refer to such sisters as `redundant'. To identify redundancy, we defined a similarity metric (see \emph{Methods} section and Figure~\ref{fig:simdiagram}). Our similarity metric can be used to identify sisters that have very high similarity between them; alternatively, redundant sisters could be identified if a single sister had a very high similarity score. We referred to these two scoring criteria as `avg' and `any' respectively.

\subsubsection*{Terminal branch terms and parent terms can be safely removed in an algorithmic fashion }
Another problem arises from the fact that the tissue ontology is scarcely populated at this point in time. Many nodes have 0-10 annotations, which we consider too few to accurately test. To solve this issue, we implemented a straightforward trimming algorithm. For a given terminal node, we test whether the node has more than a threshold number of annotations. If it does not, the node is removed. The next node in the branch is tested and removed recursively until a node which satisfies the condition is found. At that point, no more nodes can be removed from that branch. This is guaranteed by the structure of the ontology: Parent nodes inherit all of the annotations of all of their descendants, so the number of annotated terms monotonically increases with increasing term hierarchy (see Figure~\ref{fig:trim_ends}). In this way, we ensure that our term dictionary includes only those tissues that are considered sufficiently well annotated for statistical purposes.

@@

\subsection*{Validation of the algorithm and parameter selection}
	We wanted to select a dictionary which included enough terms to be specific beyond the largest  \emph{C. elegans}tissues, yet would minimize the number of spurious results and which had a good dynamic range in terms of enrichment fold-change. Selection of a dictionary based only on minimization of spurious results would result in a dictionary with a large number of annotations per tissue, and would therefore include only the major tissues. On the other hand, selecting a dictionary that can detect smaller tissues will bias us towards tissues with lesser annotations. To our knowledge there is no good method for assessing false-positive or false-negative results for annotations.To help us select an appropriate dictionary and validate our tool, we found a set of 30 gold standards based on microarray and RNA-seq literature which are believed to be enriched in specific tissues \cite{Gaudet2004a, Spencer2011, Cinar2005, Watson2008a, Pauli2006, Portman2004, Fox2007, Smith2010}. Some of these studies have been used to annotate gene expression, and so they did not constitute an independent testing set. To correct this flaw, we built a clean dictionary that specifically excluded all annotation evidence that came from these studies.

	As a first attempt to select a good dictionary, we generated all the possible combinations of dictionaries with minimal annotations of 10, 25, 50 and 100 genes and similarity cutoffs of 0.9, 0.95 and 1, using `avg' or `any' thresholding criteria for the latter (see Table ~\ref{tab:DictionarySpecs}). For these dictionaries, the number of tissues tested ranged from 21 to 460. The number of tissues was inversely correlated to the minimum annotation, as expected, and was largely insensitive to the redundancy threshold, at least in the range we explored (0.9-1). Next, we analyzed all 30 datasets using each dictionary. Because of the large number of results, instead of analyzing each set of terms individually, we pooled all results for a given dictionary into histograms. When we analyzed the distribution of significant q-values for the dictionaries, we found that the similarity threshold mattered relatively little for any dictionary. We also noticed that the `any' thresholding method resulted in tighter histograms with a mode closer to 0. For this reason, we chose the `any' method for dictionary generation. The average q-value increased with decreasing annotation cut-off (see Figure~\ref{fig:qvals}), which reflects the decreasing statistical power associated with fewer annotations per term, but we remained agnostic as to how significant is the trade-off between power and term specificity. Based on these observations, we ruled out the dictionary with the 100 gene annotation cut-off: it had the fewest terms and its q-values were not low enough to compensate the trade-off in specificity.

@@

\subsection*{A brief example}
We applied our tool to the RNA-seq datasets developed by Engelmann et al.  \cite{Engelmann2011} to gain further understanding of the biology underlying these datasets. Engelmann et al.  exposed young adult worms to 5 different pathogenic bacteria or fungi for 24 hours, after which mRNA was extracted from the worms for sequencing. We ran TEA on the genes Engelmann \emph{et al} identified as up- or down-regulated. Initially we noticed that genes that are down-regulated tend to be twice better annotated on average than genes that were up-regulated, suggesting that our understanding of the worm immune system is scarce, in spite of important advances made over the last decade. Strikingly, three out of the five samples showed enrichment of neuronal tissues or neuronal precursor tissues amongst the down-regulated genes (see \ref{fig:Dcon, fig:Hsp, fig:Smar}). A possible explanation for this might be that the infected worms are sick and the neurons are beginning to shut down; an alternative hypothesis would be that the worm is down-regulating specific neuronal pathways as a behavioral response against the pathogen. Indeed, several studies \cite{Meisel2014, Zhang2005} have provided evidence that   \emph{C. elegans} uses chemosensory neurons to identify pathogens. Interestingly, one bacterium did not exhibit the same pattern of down-regulation of neuronal-associated genes. \emph{E. faecalis} showed increased expression of genes associated with neuronal tissues, hinting that \emph{E. faecalis} may have a different pathogenic profile. Up-regulated tissues, when detected, almost always included the hypodermis and excretory duct.
	Our results highlight the involvement of various   \emph{C. elegans} neuronal tissues in pathogen defense and/or illness.

\section*{Discussion}
We have presented a tissue enrichment analysis tool that employs a standard hypergeometric model to test the   \emph{C. elegans} tissue ontology. We have also presented the first, to our knowledge, ontology trimming algorithm for biomedical ontologies. This algorithm, which is very easy to execute, places strong limits on the number of terms selected for testing. Due to the nature of all ontologies as hierarchical, acyclical graphs with term inheritance, term annotations are correlated along any given branch. This correlation reduces the benefits of including all terms for statistical analysis: for any given term along a branch, if that term passes significance, there is a high probability that many other terms along that branch will also pass significant. If the branch is enriched by random chance, error propagation along a branch means that many more false positives will follow. Thus, a researcher might be misled by the number of terms of correlated function and assign importance to this finding; the fact that the branching structure of GO amplifies false positive signals is a powerful argument for either reducing branch length or branch intracorrelation, or both. On the other hand, if a term is actually enriched, we argue that there is little benefit to presenting the user with additional terms along that branch. Instead, a user will benefit most from testing sparsely along the tree at a suitable specificity for hypothesis formation. Related terms of the same level should only be tested when there is sufficient annotation to differentiate, with statistical confidence, whether one term is enriched above the other. Our algorithm reduces branch length by identifying and removing nodes that are insufficiently annotated and parents that are likely to include sparse information.

Chikina \emph{et al}  \cite{Chikina2009} report a tissue enrichment model based on a Support Vector Machine classifier that has been trained on microarray studies. SVM classifiers are powerful tools capable of great sensitivity, but they require continuous retraining as tissue expression data becomes more available. Moreover, classifiers require that data be rank-ordered by some metric, something which is not possible for some genome-wide studies. Our tool relies on an annotation dictionary that is continuously updated, does not require retraining and does not require ranked genes. To our knowledge, there are no other tissue ontology enrichment tools in   \emph{C. elegans}, but similar projects exist for humans and zebrafish ~\cite{Lee2013, Prykhozhij2013}, highlighting the relevance of our tool for high-dimensionality biology.

@@

\section*{Availability of data and materials}
Our web implementation is available at https://www.WormBase.org/tea. Our software can also be downloaded using Python's pip installer via the command

@@

\section*{Competing interests}
 	The authors declare that they have no competing interests.
\section*{Author's contributions}
    DA and PWS conceived of the project; DA developed algorithm;
	RYL made intellectual contributions to the project; RYL and JC
	developed the web GUI.
\section*{Acknowledgements}
	We thank Justin Bois for his help and support. 
	We would like to acknowledge all members of the Sternberg
	lab for helpful discussion.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
